Using Genetic Algorithms to Find Approximations for the Minimum Vertex Cover Problem

### 2018-02-28 paper.tex
reduced margins
done

% including packages
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[autostyle]{csquotes}
\usepackage{biblatex}
\usepackage{amsfonts}
\usepackage[end]{algpseudocode}
\usepackage{caption}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage[a4paper, total={6in, 8in}]{geometry}
\usepackage{algorithm}
\usepackage[all]{hypcap}

@@

% table of contents page
\setstretch{0.5}
\phantomsection\addcontentsline{toc}{section}{Contents}
\tableofcontents
\setstretch{1.0}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\textwidth]{ga}
\caption{General schema for genetic algorithms}
\label{fig:ga}
\end{figure}

@@

\begin{algorithm}[!htbp]
\caption{GenticAlgorithm$\left( n, u, p_c, p_m, f \right)$}
\label{alg:ga}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.5\textwidth]{ps}
\caption[The regular graph after Papadimitriou
and Steiglitz for $k=3$]
{Construction of the regular graph after Papadimitriou
and Steiglitz with $k=3$}
\label{fig:ps}
\end{figure}

Fortunately, a simpler and better algorithm exists.
The following simple algorithm is surprisingly the
best approximation algorithm known for the mvcp \cite[p.~301]{9:mvcp}.

\begin{algorithm}[!htbp]
\caption{vercov($G = (V, E)$)}
\label{alg:vercov}
\begin{algorithmic}[1]

@@

\begin{algorithm}[!htbp]
\caption{GenerateRandomGraph$(n, d, k)$}
\label{alg:gen}
\begin{algorithmic}[1]
\algnotext{EndFor}
\algnotext{EndIf}

@@

\begin{table}[!htbp]
\centering
\includegraphics[width=1\textwidth]{t3}
\caption[Results obtained by GA for graphs in test set~2]{%
Experimental results obtained by the genetic algorithm for five random graphs of size $n = 200$ with edge density: $d = 0.1$ (``mvcp200-01''), $d = 0.2$ (``mvcp200-02''), $d = 0.3$ (``mvcp200-03''), $d = 0.5$ (``mvcp200-04''), $d = 0.5$ (``mvcp200-05'') and the regular graph of size $n = 202 \ (k=66)$ from Papadimitriou and Steiglitz (``PS202'').%
}
\label{tbl:t3}
\end{table}

\begin{table}[!htbp]
\centering
\includegraphics[width=1\textwidth]{t4}
\caption[Results obtained by vercov heuristic for graphs in test set~2]{%
Experimental results obtained by the vercov heuristic for five random graphs of size $n = 200$ with edge density: $d = 0.1$ (``mvcp200-01''), $d = 0.2$ (``mvcp200-02''), $d = 0.3$ (``mvcp200-03''), $d = 0.5$ (``mvcp200-04''), $d = 0.5$ (``mvcp200-05'') and the regular graph of size $n = 202 \ (k=66)$ from Papadimitriou and Steiglitz (``PS202'').%
}
\label{tbl:t4}
\end{table}

@@

The average performance of the vercov heuristic (surprisingly)
decreases as the edge density is increased;
i.e. the problems become even harder for the vercov heuristic
while they become much simpler for the genetic algorithm when
the edge density increases.
The average fitness found by the genetic algorithm is notably
better than the one found by the vercov heuristic.
One can demonstrate this by calculating
the ratio of the absolute (or equivalently, the relative) error.
For example, for the first problem instance,
let $\bar{f}_1, \bar{f}_2$ be the average found by
the genetic algorithm, vercov heuristic respectively.
We have $ (\bar{f}_2 - 55) / (\bar{f}_1 - 55) = (86.46 - 55)/(62.61 - 55) \approx 4 = 400\%$
that is, the error was reduced by a factor of 4.

@@

>>>>

% including packages
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[autostyle]{csquotes}
\usepackage{biblatex}
\usepackage{amsfonts}
\usepackage[end]{algpseudocode}
\usepackage{caption}
\usepackage{setspace}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage[a4paper, total={6in, 8.5in}]{geometry}
\usepackage{algorithm}
\usepackage[all]{hypcap}

@@

% table of contents page
\phantomsection\addcontentsline{toc}{section}{Contents}
\tableofcontents

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=1\textwidth]{ga}
\caption{General schema for genetic algorithms}
\label{fig:ga}
\end{figure}

@@

\begin{algorithm}[!htb]
\caption{GenticAlgorithm$\left( n, u, p_c, p_m, f \right)$}
\label{alg:ga}

@@

\begin{figure}[!htb]
\centering
\includegraphics[width=0.5\textwidth]{ps}
\caption[The regular graph after Papadimitriou
and Steiglitz for $k=3$]
{Construction of the regular graph after Papadimitriou
and Steiglitz with $k=3$}
\label{fig:ps}
\end{figure}

Fortunately, a simpler and better algorithm exists.
The following simple algorithm~\ref{alg:vercov} is surprisingly the
best approximation algorithm known for the mvcp \cite[p.~301]{9:mvcp}.

\begin{algorithm}[!htb]
\caption{vercov($G = (V, E)$)}
\label{alg:vercov}
\begin{algorithmic}[1]

@@

\begin{algorithm}[!htb]
\caption{GenerateRandomGraph$(n, d, k)$}
\label{alg:gen}
\begin{algorithmic}[1]
\algnotext{EndFor}
\algnotext{EndIf}

@@

\begin{table}[!htb]
\centering
\includegraphics[width=1\textwidth]{t3}
\caption[Results obtained by GA for graphs in test set~2]{%
Experimental results obtained by the genetic algorithm for five random graphs of size $n = 200$ with edge density: $d = 0.1$ (``mvcp200-01''), $d = 0.2$ (``mvcp200-02''), $d = 0.3$ (``mvcp200-03''), $d = 0.5$ (``mvcp200-04''), $d = 0.5$ (``mvcp200-05'') and the regular graph of size $n = 202 \ (k=66)$ from Papadimitriou and Steiglitz (``PS202'').%
}
\label{tbl:t3}
\end{table}

\begin{table}[!htb]
\centering
\includegraphics[width=1\textwidth]{t4}
\caption[Results obtained by vercov heuristic for graphs in test set~2]{%
Experimental results obtained by the vercov heuristic for five random graphs of size $n = 200$ with edge density: $d = 0.1$ (``mvcp200-01''), $d = 0.2$ (``mvcp200-02''), $d = 0.3$ (``mvcp200-03''), $d = 0.5$ (``mvcp200-04''), $d = 0.5$ (``mvcp200-05'') and the regular graph of size $n = 202 \ (k=66)$ from Papadimitriou and Steiglitz (``PS202'').%
}
\label{tbl:t4}
\end{table}

@@

The average performance of the vercov heuristic (surprisingly)
decreases as the edge density is increased;
i.e. the problems become even harder for the vercov heuristic
while they become much simpler for the genetic algorithm when
the edge density increases.
The average fitness found by the genetic algorithm is notably
better than the one found by the vercov heuristic.
One can demonstrate this by calculating
the ratio of the absolute (or equivalently, the relative) error.
For example, for the first problem instance,
let $\bar{f}_1, \bar{f}_2$ be the average found by
the genetic algorithm, vercov heuristic respectively.
We have $ (\bar{f}_2 - 55) / (\bar{f}_1 - 55) = (86.46 - 55)/(62.61 - 55) \approx 4 = 400\%$
that is, the error was reduced by a factor of 4.
This is further visualized in figure~\ref{fig:av} for the random graphs in
test set~1 (figure on the left) and for the random graphs in test set~2 (figure on the right),
where the average value found by each algorithm
for each graph instance is plotted.

\begin{figure}[!h]

\begin{subfigure}{0.5\textwidth}
\includegraphics[width=1\linewidth]{av1} 
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=1\linewidth]{av2}
\end{subfigure}

\caption[average fitness value plotted against edge density]
{%
average fitness value found by GA and vercov heuristic plotted
against the edge density for each random graph in test set~1 (the left figure) or~2
(the right figure).%
}
\label{fig:av}
\end{figure}

@@

%\pagebreak
